Module 5 Project 2: Storm Data Analysis
=======================================

## Synopsis
The report presents the results of an analysis on storm data collected by the U.S. National Oceanic and Atmospheric Administration.

Usingthe data, the following types of events were determined: 
* cause greatest harm to the population health, and 
* have the greatest economic damages.

The purpose of the report serve to warn the government and the muincipal managers ahead of a disaster so that preparatory work could be planned ahead to reduce potental damages to the community and the economy to the minimum.


## Data Processing 

### Step 1: Dataset source and loading

The storm dataset source is 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2'.

```{r echo=TRUE, cache=TRUE}
storm_dataset<-read.csv ("repdata_data_stormData.csv")

names (storm_dataset)

```

### Step 2: Data analysis Part 1

To determine the event type that cause greatest harm to the population health, I
looked at the 'FATALITIES' and 'INJURIES' data fields. 

First, I created a subset of data with 'EVTYPE', 'FATALITIES' and 'INJURIES' data fields. Then I summed up the fatalities and injuries for each event. Those events with no fatalities or injuries were dropped. Lastly, I aggregated the combined fatalities and injuries by the event type.

```{r echo=TRUE}
x<-storm_dataset[,c("EVTYPE", "FATALITIES", "INJURIES")]

x[,"COMBINED"]<-x[,"FATALITIES"]+x[,"INJURIES"]

x <- x[x$COMBINED>0,]

library (dplyr)

x<-group_by (x, EVTYPE)

event_publichealth<-summarise(x, total=sum(COMBINED))

summary (event_publichealth)
```

I looked at summary the dataset 'event_publichealth' and there were many events with low fatalities or injuries. The mean was 707.60 while the median was 5.00. I decided to drop those events with fatalities or injuries below the mean value.

```{r echo=TRUE}
event_publichealth<-event_publichealth[event_publichealth$total>mean(event_publichealth$total),]
```

```{r echo=TRUE}
library (ggplot2)

output_plot <- qplot (x=log10(total), y=EVTYPE, data=event_publichealth)

output_plot + geom_line() + labs (y = "Event Type") + labs (x = "(No. of fatalities and/or injuries) - Log10 ") + labs (title = "Storm Event versus Fatality/Injury Count")

```

Clearly from the chart, TORNADO caused the most fatalities and injuries.

### Step 3: Data analysis Part 2

To determine the event type that cause greatest economic damages, I
looked at the 'PROPDMG' and 'CROPDMG' data fields. 

First, I created another subset of data with 'EVTYPE', 'PROPDMG' and 'CROPDMG' data fields. Then I summed up the damages for each event. Those events with no damages were dropped. Lastly, I aggregated the total damages by the event type.

```{r echo=TRUE}
y<-storm_dataset[,c("EVTYPE", "PROPDMG", "CROPDMG")]

y[,"TOTAL"]<-y[,"PROPDMG"]+y[,"CROPDMG"]

y <- y[y$TOTAL>0,]

y<-group_by (y, EVTYPE)

event_damage<-summarise(y, total=sum(TOTAL))

summary (event_damage)
```

I looked at summary the dataset 'event_damage' and there were many events with low damages. The mean was 28,451 while the median was only 59. I decided to drop those events with damages below the mean value.

```{r echo=TRUE}
event_damage<-event_damage[event_damage$total>mean(event_damage$total),]
```

```{r echo=TRUE}
output_plot <- qplot (x=log10(total), y=EVTYPE, data=event_damage)

output_plot + geom_line() + labs (y = "Event Type") + labs (x = "(Damages in dollars) - Log10 ") + labs (title = "Storm Event versus Damages")

```

Clearly from the chart, TORNADO also caused the most economic damages.

## Results

From the analyses above, I concluded that TORNADO caused the most harm to public health as well as the most economic damages.


